---
title: "epi"
author: "studentar"
date: "January 25, 2018"
output:
  html_document: default
  pdf_document: default
---
# Introduction
This is an exploratory look at the Environmental Performance Index.

# Load packages
```{r message = FALSE, warning = FALSE}
require(dplyr)
require(plotly)
```

# Read in data
There were 180 countries in 40 categories reduced to 10 categories for exploratory purposes.

```{r, include = FALSE} 
setwd("E:/epi/data")
epi <- readxl::read_xlsx("2016-epi.xlsx", sheet = 3)
```
```{r, results = 'hide'}
dim(epi)
```

# Data Preprocessing
```{r results = 'hide'}
names(epi)[names(epi) == 'Country'] <- 'country'
names(epi)[names(epi) == '2016 EPI Score'] <- 'epi_score'
names(epi)[names(epi) == 'Environmental Health'] <- 'environmental_health'
names(epi)[names(epi) == 'EH - Health Impacts'] <- 'health_impacts'
names(epi)[names(epi) == 'EH - Air Quality'] <- 'air_quality'
names(epi)[names(epi) == 'EV - Water Resources']<- 'water_resources'
names(epi)[names(epi) == 'Household Air Quality'] <-  'household_air_quality'
names(epi)[names(epi) == 'Access to Sanitation'] <-  'access_to_sanitation'
names(epi)[names(epi) == 'Access to Drinking Water'] <-  'access_to_drinking_water'
names(epi)[names(epi) == 'Wastewater Treatment'] <-   'wastewater_treatment'
```

```{r echo = FALSE}
epi <- epi[c(-1,-3,-5,-7,-10,-12,-13,-14,-15,-16,-17,-19,-20,-21,-22,-23,-25,-27,-29,-30,-31,-32,-33,-34,-35,-36,-37,-38,-39,-40)]
```

# Structure of the data
```{r results = 'hide'}
str(epi)
```

# Summary of the data
```{r results = 'hide'}
summary(epi)
```

# Results
Scatterplots and boxplots were used to explore and visualize the data.

## Health and Environmental Impact Plots
As air quality increases there isa positive impact on health and the environment in countries around the world.

```{r include = FALSE}
epi.plot = ggplot(epi, aes(x = air_quality, y = health_impacts)) +
geom_point() + geom_smooth(method="lm", color = "red")
```
```{r results = 'hide'}
epi.plot
```

```{r include = FALSE}
epi.plot = ggplot(epi, aes(x = air_quality, 
y = environmental_health)) + geom_point() + 
geom_smooth(method="lm", color = "blue")
```
```{r results = 'hide'}
epi.plot
```

```{r echo = FALSE}
epi_air = sample(epi$air_quality, replace = T)
epi_environmental = sample(epi$environmental_health, replace = T)
C = list(epi_air, epi_environmental)
names(C) = c(paste("Air Quality\n n = " , length(epi_air) , 
sep=""), paste("Environmental Health\n n = " , 
length(epi_environmental) , sep=""))
par(mgp=c(3,2,0))
boxplot(C, col = rgb(0.9,0.9,0.9,0.9), ylab = "value" )
```
```{r echo = FALSE}
epi_air = sample(epi$air_quality, replace = T)
epi_health = sample(epi$health_impacts, replace = T)
C = list(epi_air, epi_health)
names(C) = c(paste("Air Quality\n n = " , length(epi_air) , 
sep=""), paste("Health Impacts\n n = " , 
length(epi_health) , sep=""))
par(mgp=c(3,2,0))
boxplot(C, col = rgb(0.9,0.9,0.9,0.9), ylab = "value" )
```

# Conclusion
This is an exploratory look at the Environmental Performance Index on air quality, the environment and health impacts.